Drivers of sex differences in consumption

### 2017-07-05 sex-difference/sex-difference.Rmd
clean up sex

```{r load_data}
sum_over = function(target_columns) {
  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
  dummy_vars_string = paste0(dummy_vars, collapse=',')
  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
  lazyeval::interp(f, .values=dummy_vars_list)
}

@@

Does that mean that:

- women are older and therefore use more azithromycin,
- women have more chronic conditions (esp. COPD) and therefore use more, or
- women have more encounters with doctors and therefore are prescribed more?

### Encounters by sex

Do women have more encounters than men? Does that explain their increased usage?

```{r load_carrier_data, eval=FALSE}
carrier = read_tsv('../../data/carrier_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_carrier_dates=n_from_dates)

inpatient = read_tsv('../../data/inpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_inpatient_dates=n_from_dates)

outpatient = read_tsv('../../data/outpatient_count_2011.tsv') %>%
  rename(bene_id=BENE_ID, n_outpatient_dates=n_from_dates)

denom = usage %>% select(bene_id, sex, age) %>% distinct
```

```{r azithromycin_usage, eval=FALSE}
na_to_zero = function(x) if_else(is.na(x), 0L, x)

azi_usage = usage %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>%
  summarize(n_claims=sum(n_claims)) %>%
  left_join(denom, by='bene_id') %>%
  filter(age >= 65, sex %in% c('male', 'female')) %>%
  left_join(carrier, by='bene_id') %>%
  left_join(inpatient, by='bene_id') %>%
  left_join(outpatient, by='bene_id') %>%
  mutate(age_group=bounds_group(age, c(65, 75, 85, 95))) %>%
  mutate_at(vars(n_claims, n_carrier_dates, n_inpatient_dates, n_outpatient_dates), na_to_zero)
```

```{r azithromycin_usage_table, eval=FALSE}
azi_usage %>%
  group_by(age_group, sex) %>%
  summarize(n_beneficiaries=n(),
            mean_n_azithromycin_claims=mean(n_claims),
            med_n_carrier_dates=median(n_carrier_dates),
            med_n_outpatient_dates=median(n_outpatient_dates),
            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
  kable
```

And I'll do a Poisson regression of number of azithromycin claims against age, being female, and number of carrier dates:

```{r azithromycin_usage_model, eval=FALSE}
library(broom)
azi_usage %>%
  mutate(is_female=sex=='female') %>%
  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
  tidy %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
  kable
```

The coefficient for being female (0.05) accounts for almost the entire difference in consumption between men and women (56 per thousand). Note that age has an approximately zero correlation (i.e.,

@@

```{r consumption_by_age_and_sex, eval=FALSE}
bene %>%
  left_join(total_pde, by='bene_id') %>%
  replace_na(list(n_claims=0)) %>%
  group_by(sex, age) %>%
  summarize(mean_n_claims=mean(n_claims)) %>%
  ungroup() %>%
  ggplot(aes(x=age, y=mean_n_claims, color=sex)) +
  geom_point()
```

@@

The older people are more female: age 65 is about 60% female. This stays constant around 60% (or dips slightly) until age 75, when it steadily increases up to 85-90% female at age 100. So we'll need to account for this age difference.

```{r sex_by_age, eval=FALSE}
plot_dat = usage %>%
  select(age, sex) %>%
  filter(sex %in% c('male', 'female')) %>%
  group_by(age) %>% count(sex) %>% ungroup() %>%
  spread(sex, n, fill=0L) %>%
  mutate(total=female+male, grand_total=sum(total),
         fraction_year=total/grand_total, fraction_female=female/total)

ggplot(mapping=aes(x=age, y=fraction)) +
  geom_point(data=rename(plot_dat, fraction=fraction_female)) +
  geom_line(data=mutate(plot_dat, fraction=fraction_year*10), col='red') +
  ylab('fraction female') + ylim(0, 1) +
  theme_minimal()
```

@@

Maybe women take more azithromycin because usage is more concentrated in women than in men. Then we could explain the increased consumption by a small number of people taking more drug. (Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women. As always, this is driven by the extensive fraction: more women took at least one dose:

```{r, eval=FALSE}
pde %>%
  filter(drug=='azithromycin') %>%
  count(bene_id) %>% rename(n=n_azi) %>%
  right_join(bene, by='bene_id') %>% replace_na(list(n_azi=0)) %>%
  count(sex, n_azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(n_azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```

>>>>

```{r load_data, cache=FALSE}
sum_over = function(target_columns) {
  dummy_vars = sapply(1:length(target_columns), function(i) sprintf('x%i', i))
  dummy_vars_string = paste0(dummy_vars, collapse=',')
  dummy_vars_list = lapply(target_columns, as.name) %>% setNames(dummy_vars)
  f = as.formula(paste0(c('~sum(', dummy_vars_string, ')'), collapse=''))
  lazyeval::interp(f, .values=dummy_vars_list)
}

@@

@@

```{r consumption_by_age_and_sex}
bene %>%
  filter(in_cohort) %>%
  group_by(bene_id) %>%
  summarize(is_female=any(is_female),
            n=sum(n_claims),
            age=min(age)) %>%
  ggplot(aes(x=age, y=n, group=is_female)) +
  geom_boxplot()
```

@@

The older people are more female.

```{r sex_by_age}
bene %>% filter(in_cohort) %>%
  ggplot(aes(x=age, fill=sex)) +
  geom_bar(position='dodge')
```

@@

Maybe women take more azithromycin because usage is more concentrated in women than in men. Then we could explain the increased consumption by a small number of people taking more drug. (Yonatan points to Lady Windmere syndrome.) In fact, consumption is more even among women; this is driven by the extensive fraction: a greater fraction of women (than men) took at least one dose.

```{r azithro_even}
pde %>%
  filter(antibiotic=='azithromycin') %>%
  group_by(bene_id) %>% summarize(azi=sum(n_claims)) %>%
  right_join(bene %>% filter(in_cohort) %>% select(bene_id, sex) %>% distinct,
    by='bene_id') %>% replace_na(list(azi=0)) %>%
  count(sex, azi) %>%
  group_by(sex) %>%
  mutate(f_ppl=n/sum(n), cum_f_ppl=cumsum(f_ppl),
         cum_cons=cumsum(azi*n), cum_f_cons=cum_cons/max(cum_cons)) %>%
  ggplot(aes(x=cum_f_ppl, y=cum_f_cons, color=sex)) +
  geom_point() +
  geom_line() +
  xlab('Fraction of consumers') + ylab("Fraction of that sex's azithromycin consumed")
```
